---
title: "P8105 Homework 5"
author: "Maria Serafini"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
  )

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
  )

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

The code below creates a function that, for a fixed group size, randomly draws "birthdays" for each person, checks whether there are duplicate birthdays in the group, and returns `TRUE` or `FALSE` based on the result.

```{r}
bday_sim = function(n_room) {
  
  birthdays = sample(1:365, n_room, replace = TRUE)

  repeated_bday = length(unique(birthdays)) < n_room

  repeated_bday
  
}
```


The code below runs the `bday_sim` function 10,000 times for each group size between 2 and 50, and computes the probability that at least two people in a given group will share a birthday. 

```{r}
bday_sim_results <- 
  expand_grid(
    bdays = 2:50,
    iter = 1:10000
  ) |> 
  mutate(
    result = map_lgl(bdays, bday_sim)
  ) |> 
  group_by(
    bdays
  ) |> 
  summarize(
    prob_repeat = mean(result)
  ) |> 
  view()
```


This plot shows the probability calculated above as a function of group size. 

```{r}
bday_sim_results |> 
  ggplot(aes(x = bdays, y = prob_repeat)) +
  geom_point() +
  geom_line()
```

From the plot we can see that as group size increases, the probability that at least two people will share a birthday increases. There is a 25% chance of at least two people sharing a birthday once the group size reaches 15, a 50% chance once the group size reaches 23, a 75% chance once the group size reaches approximately 32, and almost a 100% chance once the group passes 50. These results indicate that a large sample size is not necessarily needed for two people to likely share a birthday. 